### 2021-04-21 assertions/doc/thesis-bernhardt/chapters/2-use_case_analysis.tex
Own chapt probabilistic model

\chapter{Use-case analysis}\label{chap:use_cases}
The introduction argued that validating properties on input parameters becomes expensive when a data structure (such as arrays), certain data types (such as byte sequences) or a range of values have to be iterated. It showed examples for such properties and identified logical formulas containing universal quantification as amenable for distributed assertion checking. Through its domain of discourse, the quantification spans an iteration space whose elements can be checked independently by each validator. As a matter of fact, the same is also valid for existential quantification, even though its negation has a different notion than that of its universal counterpart. The difference between the two is addressed later in this chapter. \\
The following section formally defines a set of applicable logical formulas for the basic implementation of this approach. Given this set, section \secref{sec:examples} gives some more in-depth examples of such formulas. As a benchmark for the distributed scheme, the last section formalizes the cost incurred by checking formulas on Tezos locally.

\section{Definition of a set of logical formulas}\label{sec:formulae}
The set of formulas that well-formed assertions can express is given in \figref{fig:formulas}. This corresponds, with some restrictions, to the language of second-order predicate logic. Besides universal and existential quantifiers with exactly one bound variable, the set definition contains logical connectives, relational symbols, constant values of a numerical, string or boolean type, variables and the basic functional symbols for basic arithmetic operations. This set of symbols is rather generic and should be valid for any target blockchain. Specific to the target blockchain are the sets of types $t$ and functional symbols $f$. These can only contain types and operations that can be translated to operation codes (opcodes) of the respective Smart Contract language and executed within the blockchain's virtual machine (VM). As an example, the primitive types in Michelson are \texttt{string}, \texttt{nat}, \texttt{int}, \texttt{bytes}, whereas Ethereum's Smart Contract language Solidity  \cite{solidity_docs} additionally supports an \texttt{enum} type and fixed-point numbers.

\begin{align*}
    t &::= \text{primitive data types (existing in the blockchain VM)} \\
    \sigma &::= \forall \mid \exists \\
    \Theta &::= \Phi \mid \sigma (x:t) \Theta \\
    \Phi,\Psi &::= \neg\Phi \mid \Phi \Rightarrow \Psi \mid \Phi\wedge\Psi \mid
    				\Phi\vee\Psi \mid \Phi \oplus \Psi \mid M \rho N \\
    \rho &::= < \mid > \mid \le \mid \ge \mid = \mid \ne \\
    M, N &::= x \mid c \mid M \odot N  \mid f (\overline M) \\
    \odot &::= +\mid -\mid * \mid / \mid \% \\
    c &::= \text{numerical value} \mid string \mid \top \mid \bot \\
    f &::= \text{operations (existing in the blockchain VM)}
\end{align*}
\begingroup\vspace*{-\baselineskip}
\captionof{figure}{Set of logical formulas amenable to distributed assertion checking (based on \cite{thiemann_2020})}
\label{fig:formulas}
\vspace*{\baselineskip}\endgroup

@@

\subsection{Two numbers are coprime}\label{sec:coprime}
The greatest common divisor ($gcd$) of two non-zero integers $a$ and $b$ is the largest integer that divides both evenly \cite{hardy2008introduction}. A more specific $gcd$-problem is finding two numbers which are coprime, meaning their greatest common divisor is one, i.e., $gcd(a, b) = 1$. In order to verify that two numbers are coprime, it has to be checked that there is no $gcd > 1$, hence the property of coprimeness can be expressed with the following formula in predicate logic:
\begin{equation}\label{eq:coprime-universial}
    (\forall n : int) (2 \le n \le \min(a,b)) \Rightarrow \neg((a \mathbin{\%} n) = 0 \land (b \mathbin{\%} n) = 0)
\end{equation}
Assuming constant time for calculating the two remainders, checking whether two numbers satisfy this property takes linear time, i.e. $\mathcal{O}(n)$, and depends on the size of the smaller number.

@@

\subsection{Existential quantification - intersection of sets}\label{sec:existential}
So far, this thesis only considered formulas featuring universal quantification, for which the assertion scheme challenged the validators to find counterexamples. However, distributed assertion checking can not only be used to refute assumptions, but also to prove their validity. This is the case for formulas featuring existential quantification. As an example, consider the assumption that two sets $U$ and $V$ have a non-empty intersection, i.e., $U \cap V \neq \emptyset$. This assumption and its negation can be expressed in predicate logic by 
\begin{equation}\label{eq:intersect}
  (\exists i, \exists j) (0 \le i < |u|) \land (0 \le j < |v|) \Rightarrow u[i] = v[j]
\end{equation}
and
\begin{equation}\label{eq:intersect_neg}
  (\forall i, \forall j) (0 \le i < |u|) \land (0 \le j < |v|) \land u[i] \neq v[j]
\end{equation}
If the condition $u[i] \neq v[j]$ is false for random values of $i$ and $j$, it does not produce a counterexample --- it produces a proof for the validity of the assumption, because it identified an element that is present in both sets. In that case, the validator finding such a proof should publish an approval instead of a veto to the network. After an approval has been posted, and all validators have come to a consensus about the validity of the assumption, the transaction may be included in a block (assuming it doesn't fail due to other reasons). For the blockchain protocol, this implies that distributed refutation and validation have to be implemented as separate modi of assertion checking. \\

@@

\subsection{Nested quantification - SAT problems}
The previous example showed that universal and existential quantification have to be handled differently. However, further special cases caused by nested quantification have to be considered as well. A nesting of the same type of quantifier, e.g. $\forall i \forall j$ can be considered as trivial. They can be handled by generating a random value for each predicate variable and checking for a counterexample or proof respectively. It becomes less trivial though, if the nesting contains both universal and existential quantifiers. This section examines the cases of $\forall i \exists j$ and $\exists i \forall j$ on the basis of a contract that expects a solution for a boolean satisfiability problem (SAT). \\

@@

\subsubsection{Encoding of variables, literals and clauses}
A SAT problem can be represented as a list of clauses $c$ and a variable assignment $\mathcal{A}$. To avoid having to handle the variables and their literals as strings and thus optimize runtime and memory consumption, an encoding scheme taken from this blogpost \cite{sabablog} is used: \\
Variables are encoded by a unique number starting from $0$ to $n-1$, where $n$ is the total number of variables. The positive and negative literals of a variable with id $x$ are encoded by the function
\begin{align*}
encode(x) =
\begin{cases}
  $2x$  & \text{if x is present as a positive literal}\\
  $2x+1$ & \text{if x is present as a negated literal}\\
\end{cases}   
\end{align*}
Consequently, checking whether a literal $y$ is positive or negative is a simple bit-wise $AND$ operation:
\begin{align*}
\texttt{isPositive(y) = y \& 1 == 0}
\end{align*}
Consider the following boolean formula and variable assignment:
\begin{align*}
F = (A \vee B) \wedge (B \vee \neg C) \text{, } \mathcal{A} = \{A \rightarrow false, B \rightarrow true, C \rightarrow false\}
\end{align*}
Following the encoding of variables given above, the list of variables $v=[A,B,C]$ occurring in $F$ is encoded as $v_{enc} = [0,1,2]$. If clauses are given as lists of literals, $F$ can be represented by the encoding $F_{enc} = [[0,2],[2,5]]$. Since the encoding of variables corresponds to their index in the list of variables, $\mathcal{A}$ can be passed as a list of booleans that maintains the same order of variables: $\mathcal{A}_{enc} = [False, True, False]$. If the blockchains VM supports a respective data type, $\mathcal{A}$ can alternatively be represented with a more efficient bitarray. Lastly, the variable encoding $x$ of a literal $y$ is obtained by dividing its encoding by two, which corresponds to a bit-wise shift to the right:
\begin{align*}
\texttt{decode(y) = y >> 1}
\end{align*}

@@

In the previous examples, where formulas contained only a single quantifier, the predicate variables of the negated formula are replaced by a randomly generated number during assertion checking. Quantifiers thus are translated to random generators during the generation of target code. However, by checking a random literal in a random clause, the validator cannot make a definite statement whether $\mathcal{A}$ satisfies $F$, since each clause has to be examined as a whole in order to verify that the conjunction evaluates to true. Consequently, the distribution can only take place on the level of clauses, but not on the level of literals. Each validator thus has to check all literals of one random clause, which translates into a random generator for the outer quantifier, and a loop for the inner quantifier. Whether the assertion requires a proof or a counterexample can be derived from the outer quantifier, hence validators produce a counterexample in this case. Due to the fact that the whole loop has to be run in order to identify a counterexample and the consequential code structure this translates to, the negation of the formula is not required but optional. \lstref{lst:cnf_test_code} shows the generated testing code of the CNF SAT problem with and without negating the formula first, where \texttt{check} evaluates the negated or non-negated condition respectively. While formulas with a single quantifier simply translate to generating a random value and checking the condition, this code is more extensive and requires a more sophisticated compiler that can differentiate between different quantification compositions.

\vspace{\baselineskip}
\noindent
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[label=lst:cnf_code_neg, numbers=none]
i = random(0, size(F))
for (j : [0..k)):
  if not check( ... ):
    // property holds
    return true
// counterexample found
return false
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[label=lst:cnf_code, numbers=none]
i = random(0, size(F))
for (j : [0..k)):
  if check( ... ):
    // property holds
    return true
// counterexample found
return false
\end{lstlisting}
\end{minipage}
\captionof{lstlisting}{Test code structure for single nested quantifiers with negation (left) and without negation (right)\label{lst:cnf_test_code}}
\vspace{\baselineskip}

@@

\subsubsection{Satisfiability in predicate logic - DNF}\label{sec:dnf}
The same propositional formula with the reversed quantification nesting $\exists i \forall j$ describes the assertion for the assumption $\mathcal{A}(F) = 1$, where $F$ is given in disjunctive normal form (DNF), i.e., a disjunction of conjunctions \cite{dnf_math_encycl}:
\begin{equation}\label{eq:dnf_sat}
\begin{aligned}
(\exists i : int, \forall j : int) \: &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}
The translation to code can be approached in the same way than its CNF counterpart. Instead of a counterexample, this use-case demands for a proof.

\subsection{Further nesting}
Nesting is, of course, not limited to two quantifiers only. Generally, formulas containing universal or existential quantification exclusively within the nesting benefit the most from distributed assertion checking, as each quantifier can be translated into a random generator. On the other hand, formulas with both types and a transition from one to the other early in the order benefit the least, because starting from the first transition, the subsequent quantifiers are translated to loops.\\
To substantiate this argument, consider two variations of a contract that expect a list of boolean formulas in CNF, and an assignment $\mathcal{A}$. Variation one checks whether $\mathcal{A}$ satisfies at least one $F$, the other one checks if all $F$ are satisfied. This can be expressed by nesting quantifiers as follows: 
\begin{itemize}
\item $\forall f \forall i \exists j$
\item $\exists f \forall i \exists j$
\end{itemize}
In the first case, each validator picks a random clause from a random $F$ and checks if at least one literal evaluates to true. This translates to two random generators and a loop for the existential quantifier. In the second case, checking a random clause in isolation cannot prove the satisfiability, hence the validator has to check all clauses in a loop. This translates to one random generator and two loops. The same reasoning can be done for the reverse order, i.e. $\exists\exists\forall$ and $\forall\exists\forall$.

\section{Restriction of use-cases}\label{sec:restrict}\todo{Better title? -> conclusion?}
\todo{meh}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking. Additionally,  as well as a distinct translation to target code. This increases the complexity and requirements for an implementation of the off-chain infrastructure and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

\section{Cost analysis of a local approach}\label{sec:usecase_cost}
For the purpose of having a benchmark for the evaluation of the distributed approach, this section analyses the cost of invoking a contract on the blockchain that locally checks a formula. Given that the target blockchain is Tezos, the computational and storage-related costs of a transaction are represented in units of gas. Although they are not paid directly in terms of fees, the actual fees paid to the miner (called bakers in Tezos) are based on the gas consumption. The gas model of Tezos is described in more detail in \secref{sec:tezos}. According to this model, the total gas consumption of a transaction $T(source,\, destination=C,\, arg, \, ...)$ calling a smart contract $C(storage, \, parameter\, type, \, code)$ is calculated by \eqref{eq:gas_transaction}, where $I$ is the set of internal operations:
\begin{align}\label{eq:gas_transaction}
\begin{split}
GasConsumed(T) = \quad &base \, gas \, cost \\
+& \, gasRead(code) \\
+& \, gasRead(storage) \\
+& \, gasDeserialize(storage, \, arg) \\
+& \, gasParseType(parameter \, type, \, storage) \\
+& \, gasParseData(storage, \, arg) \\
+& \, gasParseCode(code) \\
+& \, gasInterpret(code, \, arg, \, storage) \\
+& \, gasUnparse(storage) \\
+& \, gasSerialize(storage) \\
+& \, gasWrite(storage) \\
+& \, \sum_{i \in I} GasConsumed(i)
\end{split}
\end{align}
Assuming that the storage of the contract is never increased, the amount of gas returned by the cost functions independent of the argument, i.e., the parameter value, remains constant for each $T$. The cost functions for the deserialization, parsing and interpretation, however, are functions of $arg$ that grow monotonically with increasing sizes of $arg$ \cite{morley_gasmodel}\cite{tezos_repo}.

>>>>

\chapter{Use-case analysis}\label{chap:use_cases}
The introduction argued, that validating properties on input parameters can become costly when data structures (such as arrays), certain data types (such as byte sequences) or a range of values have to be iterated. It showed examples for such properties and identified logical formulas containing universal quantification as amenable for distributed assertion checking. Through its domain of discourse, the quantification spans an iteration space whose elements can be checked independently by the validators. As a matter of fact, the same is also valid for existential quantification, even though its negation has a different notion than that of its universal counterpart. The difference between the two is addressed later in this chapter.

The following section formally defines a set of applicable logical formulas for the basic implementation of the distributed assertion checking scheme. Given this set, section \secref{sec:examples} gives some more in-depth examples of such formulas. As a benchmark for the distributed scheme, the last section formalizes the cost incurred by checking formulas on Tezos locally.

\section{Definition of a set of logical formulas}\label{sec:formulae}
The set of formulas, which can be expressed in well-formed assertions, is given in \figref{fig:formulas}. This corresponds, with some restrictions, to the language of second-order predicate logic. Besides universal and existential quantifiers (with exactly one bound variable), the definition contains logical connectives, relational symbols, constant values of a numerical, string or boolean type, variables and the functional symbols for basic arithmetic operations. This set of symbols is rather generic and should be valid for any target blockchain. Specific to the target blockchain are the sets of types $t$ and functional symbols $f$. These can only contain types and operations, which can be translated to instructions of the respective Smart Contract language and interpreted by the blockchain's virtual machine (VM). As an example, the primitive types in Michelson are \texttt{string}, \texttt{nat}, \texttt{int}, \texttt{bytes}, whereas Ethereum's Smart Contract language Solidity  \cite{solidity_docs} additionally supports an \texttt{enum} type and fixed-point numbers.
\begin{align*}
    t &::= \text{primitive data types (existing in the blockchain VM)} \\
    \sigma &::= \forall \mid \exists \\
    \Theta &::= \Phi \mid \sigma (x:t) \Theta \\
    \Phi,\Psi &::= \neg\Phi \mid \Phi \Rightarrow \Psi \mid \Phi\wedge\Psi \mid
    				\Phi\vee\Psi \mid \Phi \oplus \Psi \mid M \rho N \\
    \rho &::= < \mid > \mid \le \mid \ge \mid = \mid \ne \\
    M, N &::= x \mid c \mid M \odot N  \mid f (\overline M) \\
    \odot &::= +\mid -\mid * \mid / \mid \% \\
    c &::= \text{numerical value} \mid string \mid \top \mid \bot \\
    f &::= \text{operations (existing in the blockchain VM)}
\end{align*}
\begingroup\vspace*{-\baselineskip}
\captionof{figure}{Set of logical formulas amenable to distributed assertion checking (based on \cite{thiemann_2020})}
\label{fig:formulas}
\vspace*{\baselineskip}\endgroup

@@

\subsection{Two numbers are coprime}\label{sec:coprime}
The greatest common divisor ($gcd$) of two non-zero integers $a$ and $b$ is the largest integer that divides both evenly \cite{hardy2008introduction}. A more specific $gcd$-problem is finding two numbers which are coprime, meaning their greatest common divisor is one, i.e., $gcd(a, b) = 1$. In order to verify that two numbers are coprime, it has to be verified that there is no $gcd > 1$, hence the property of coprimeness can be expressed with the following formula in predicate logic:
\begin{equation}\label{eq:coprime-universial}
    (\forall n : int) (2 \le n \le \min(a,b)) \Rightarrow \neg((a \mathbin{\%} n) = 0 \land (b \mathbin{\%} n) = 0)
\end{equation}
Assuming it takes constant time to calculate the two remainders, checking whether two numbers satisfy this property takes linear time, i.e. $\mathcal{O}(n)$, and depends on the size of the smaller number.

@@

\subsection{Existential quantification - intersection of sets}\label{sec:existential}
So far, this thesis only considered formulas featuring universal quantification, for which the assertion scheme engaged the validators to find counterexamples. However, distributed assertion checking can not only be used to refute assumptions, but also to prove their validity. This is the case for formulas featuring existential quantification. As an example, consider the assumption that two sets $U$ and $V$ have a non-empty intersection, i.e., $U \cap V \neq \emptyset$. This assumption and its negation can be expressed in predicate logic by 
\begin{equation}\label{eq:intersect}
  (\exists i, \exists j) (0 \le i < |u|) \land (0 \le j < |v|) \Rightarrow u[i] = v[j]
\end{equation}
and
\begin{equation}\label{eq:intersect_neg}
  (\forall i, \forall j) (0 \le i < |u|) \land (0 \le j < |v|) \land u[i] \neq v[j]
\end{equation}
If the condition $u[i] \neq v[j]$ is false for random values of $i$ and $j$, it does not produce a counterexample --- it produces a proof for the validity of the assumption, because it identified an element that is present in both sets. In that case, the validator finding such a proof should publish an approval instead of a veto to the network. After an approval has been posted, and all validators have come to a consensus about the validity of the proof, the result is accepted. For the blockchain protocol, this implies that distributed refutation and validation have to be implemented as separate modi of assertion checking.

@@

\subsection{Nested quantification - SAT problems}
The previous example demonstrated that universal and existential quantification have to be handled differently. However, further special cases, caused by nested quantification, have to be considered as well. A nesting of the same type of quantifier, e.g. $\forall i \forall j$, can be considered as trivial. They are translated to code, which generates a random value for each predicate variable and checks for a counterexample or proof respectively. It becomes less trivial, if the nesting contains both universal and existential quantifiers. This section examines the cases of $\forall i \exists j$ and $\exists i \forall j$ on the basis of a contract that expects a solution for a boolean satisfiability problem (SAT).

@@

\subsubsection{Encoding of variables, literals and clauses}
A SAT problem can be represented as a list of clauses $c$ and a variable assignment $\mathcal{A}$. To avoid having to handle the variables and their literals as strings, and thus optimize runtime and memory consumption, an encoding scheme \cite{sabablog} is used in the following.

Variables are encoded by a unique number starting from $0$ to $n-1$, where $n$ is the total number of variables. The positive and negative literals of a variable with id $x$ are encoded by the function
\begin{align*}
encode(x) =
\begin{cases}
  $2x$  & \text{if x is present as a positive literal}\\
  $2x+1$ & \text{if x is present as a negated literal}\\
\end{cases}   
\end{align*}
Consequently, checking whether a literal $y$ is positive or negative is a simple bit-wise $AND$ operation:
\begin{align*}
\texttt{isPositive(y) = y \& 1 == 0}
\end{align*}
Consider the following boolean formula and variable assignment:
\begin{align*}
F = (A \vee B) \wedge (B \vee \neg C) \text{, } \mathcal{A} = \{A \rightarrow false, B \rightarrow true, C \rightarrow false\}
\end{align*}
Following the encoding of variables given above, the list of variables $v=[A,B,C]$ occurring in $F$ is encoded as $v_{enc} = [0,1,2]$. If clauses are given as lists of literals, $F$ can be represented by the encoding $F_{enc} = [[0,2],[2,5]]$. Since the encoding of variables corresponds to their index in the list of variables, $\mathcal{A}$ can be passed as a list of booleans that maintains the same order of variables: $\mathcal{A}_{enc} = [False, True, False]$. If the blockchain's VM supports a respective data type, $\mathcal{A}$ can alternatively be represented with a more efficient bitarray. Lastly, the variable encoding $x$ of a literal $y$ is obtained by dividing its encoding by two, which corresponds to a bit-wise shift to the right:
\begin{align*}
\texttt{decode(y) = y >> 1}
\end{align*}

@@

In the previous examples, where formulas contained only a single quantifier, the predicate variables of the negated formula are replaced by a randomly generated number during assertion checking. Quantifiers thus are translated to random generators during the generation of target code. However, by checking a random literal in a random clause, the validator cannot make a definite statement about whether $\mathcal{A}$ satisfies $F$. Each clause has to be examined as a whole in order to verify, that the conjunction evaluates to true. Consequently, the distribution can only take place on the level of clauses, but not on the level of literals. Each validator thus has to check all literals of one random clause, which translates into a random generator for the outer quantifier, and a loop for the inner quantifier. Whether the assertion requires a proof or a counterexample can be derived from the outer quantifier, hence validators produce a counterexample in this case. \lstref{lst:cnf_test_code} shows the generated testing code of the CNF SAT problem, where \texttt{check} evaluates the negated condition. While formulas with a single quantifier simply translate to generating a random value and checking the condition, this code is more extensive. It requires a more sophisticated compiler, which can differentiate between different quantification compositions.

\begin{lstlisting}[label=lst:cnf_test_code, caption=Test code structure for single nested quantifiers, numbers=none]
i = random(0, size(F))
for (j : [0..k)):
  if not check( ... ):
    // property holds
    return true
// counterexample found
return false
\end{lstlisting}

@@

\subsubsection{Satisfiability in predicate logic - DNF}\label{sec:dnf}
The same propositional formula, with the reversed quantification nesting $\exists i \forall j$, describes the assertion for the assumption $\mathcal{A}(F) = 1$, where $F$ is given in disjunctive normal form (DNF), i.e., a disjunction of conjunctions \cite{dnf_math_encycl}:
\begin{equation}\label{eq:dnf_sat}
\begin{aligned}
(\exists i : int, \forall j : int) \: &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}
The translation to code can be approached in the same way than its CNF counterpart. Instead of a counterexample, this use-case demands for a proof.

\subsection{Further nesting}
Nesting is, of course, not limited to two quantifiers. Generally, formulas containing universal or existential quantification exclusively within the nesting benefit the most from distributed assertion checking, as each quantifier can be translated into a random generator. On the other hand, formulas with both types and a transition from one to the other early in the order benefit the least; starting from the first transition, the subsequent quantifiers are translated to loops.

To substantiate this argument, consider two variations of a contract that expect a list of boolean formulas in CNF, and an assignment $\mathcal{A}$. Variation 1 checks, whether $\mathcal{A}$ satisfies at least one $F$. Variation 2 checks, if all $F$ are satisfied. This can be expressed by nesting quantifiers as follows: 
\begin{enumerate}
\item $\forall f \forall i \exists j$
\item $\exists f \forall i \exists j$
\end{enumerate}
In the first case, each validator picks a random clause from a random $F$, and checks if at least one literal evaluates to true. This translates to two random generators and a loop for the existential quantifier. In the second case, checking a random clause in isolation cannot prove the satisfiability, hence the validator has to check all clauses in a loop. This translates to one random generator and two loops. The same reasoning can be done for the reverse order, i.e. $\exists\exists\forall$ and $\forall\exists\forall$.

\section{Restriction of use-cases}\label{sec:restrict}\todo{Better title? -> conclusion?}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking. Additionally, they require a distinct translation to target code. This increases the complexity and requirements for an implementation of the toolchain and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

\section{Cost analysis of a local approach}\label{sec:usecase_cost}
For the purpose of having a benchmark for the evaluation of the distributed approach, this section analyses the cost of invoking a contract on the blockchain, which checks a formula locally. Given that the target blockchain is Tezos, the computational and storage-related costs of a transaction are represented in units of gas. Although they are not paid directly in terms of fees, the actual fees paid to the miner (called bakers in Tezos) are based on the gas consumption. The gas model of Tezos is described in more detail in \secref{sec:tezos}. According to this model, the total gas consumption of a transaction $T(source,\, destination=C,\, arg, \, ...)$ invoking a smart contract $C(storage, \, parameter\, type, \, code)$ is calculated by \eqref{eq:gas_transaction}, where $I$ is the set of internal operations:
\begin{align}\label{eq:gas_transaction}
\begin{split}
GasConsumed(T) = \quad &base \, gas \, cost \\
+& \, gasRead(code) \\
+& \, gasRead(storage) \\
+& \, gasDeserialize(storage, \, arg) \\
+& \, gasParseType(parameter \, type, \, storage) \\
+& \, gasParseData(storage, \, arg) \\
+& \, gasParseCode(code) \\
+& \, gasInterpret(code, \, arg, \, storage) \\
+& \, gasUnparse(storage) \\
+& \, gasSerialize(storage) \\
+& \, gasWrite(storage) \\
+& \, \sum_{i \in I} GasConsumed(i)
\end{split}
\end{align}
Assuming the storage of the contract is never increased, the amount of gas returned by the cost functions independent of the argument, i.e., the parameter value, remains constant for each $T$. The cost functions for the deserialization, parsing and interpretation, however, are functions of $arg$ and grow monotonically with increasing sizes of $arg$ \cite{morley_gasmodel}\cite{tezos_repo}.
